---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(max.print = 100)
```

# R6methods

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/mattwarkentin/R6methods/workflows/R-CMD-check/badge.svg)](https://github.com/mattwarkentin/R6methods/actions)
<!-- badges: end -->

The goal of `R6methods` is to provide a _lightweight_ package that extends the S3 generic support for `R6` class objects. This package defines several S3 methods for common `R` generics (e.g. `str()`) and operators (e.g. `[` or `[<-`) to make it straightforward to define public methods in your `R6` class and have them _"just work"_.

This package is **very experimental** and liable to change drastically. Use at your own risk! Developing this package was primarily a learning experience for working with `R6` and `S3`, and may not have any practical use.

## Installation

You can install the development version of `R6methods` from [GitHub](https://github.com/mattwarkentin/R6methods) with:

``` r
remotes::install_github("mattwarkentin/R6methods")
```

## Usage

### Adding `R6methods` to your package

This package is primarily designed for use by `R` package developers. If you are developing a package which contains `R6` classes, you can save yourself extra work, such as defining `S3` methods for common `R` generics. `R6methods` is meant to be a lightweight addition for providing increased `S3` generic support.

The easiest way to benefit from this package is by depending on `R6methods` in your package `DESCRIPTION` file.
```
Package: mypackage
Title: My Package Title
Version: 0.0.0.9000
Authors@R: 
    person(given = "Jane",
           family = "Doe",
           role = c("aut", "cre"),
           email = "jane.doe@email.com")
Description: This package...
Depends:
  R6methods
```

You may optionally import specific methods using the `@importFrom` `roxygen2` tag.

### Writing dot-dunder methods

In order to benefit from the `S3` methods provided by `R6methods`, you simply need to annotate your `R6` class with so-called **dot-dunder** methods to get immediate support for many common `R` generics. They are called **dot-dunder** because the methods start with a dot (**.**) and **d**ouble-**under**score. This syntax and approach borrows inspiration from the `python` OOP. 

These dot-dunder methods must be public methods, and your class must also inherit the `R6` class (i.e. `R6::R6Class(class = TRUE)`, the default). Here is a toy example that adds support to an `R6` class `Foo` for the `[` operator.

```{r}
library(R6methods)

Foo <- R6::R6Class(
  public = list(
    x = mtcars,
    .__subset__ = function(i, j, ...) {
      self$x[i, j, ...]
    }
  )
)

foo <- Foo$new()

# Subset
foo[1:5, 1:3]
```

### Supported methods

The table below is a comprehensive list of the dot-dunder methods currently supported by `R6methods`. When creating your `R6` class, add any number of the dot-dunder methods (with the same function parameters) and gain support for the corresponding S3 method.

```{r supported-methods, echo=FALSE, includes=FALSE}
library(magrittr)
R6methods <- 
  tibble::tribble(
  ~Group, ~`Dot-Dunder Method`, ~`S3 Method`, ~Description,
  "Arithmetic" ,".__add__(e2)", "`+`", "Addition",
  "Arithmetic" ,".__sub__(e2)", "`-`", "Subtraction",
  "Arithmetic" ,".__mul__(e2)", "`*`", "Multiplication",
  "Arithmetic" ,".__div__(e2)", "`/`", "Divisiion",
  "Arithmetic" ,".__pow__(e2)", "`^`", "Powers/exponents",
  "Arithmetic" ,".__mod__(e2)", "`%%`", "Modular division",
  "Arithmetic" ,".__intdiv__(e2)", "`%/%`", "Integer division",
  
  "Subsetting" ,".__subset__(i, j, ...)", "`[`", "Subset with `[`",
  "Subsetting" ,".__subsetrep__(i, j, ..., value)", "`[<-`", "Subset assignment with `[<-`",
  "Subsetting" ,".__subset2__(i, j, ...)", "`[[`", "Subset with `[[`",
  "Subsetting" ,".__subset2rep__(i, j, ..., value)", "`[[<-`", "Subset assignment with `[[<-`",
  
  "Relational" ,".__lt__(e2)", "`<`", "Less than",
  "Relational" ,".__gt__(e2)", "`>`", "Greater than",
  "Relational" ,".__le__(e2)", "`<=`", "Less than or equal to",
  "Relational" ,".__ge__(e2)", "`>=`", "Greater than or equal to",
  "Relational" ,".__eq__(e2)", "`==`", "Equality",
  "Relational" ,".__ne__(e2)", "`!=`", "Non-equality",
  "Relational" ,".__ae__(e2)", "all.equal", "All equal",
  
  "Misc", ".__data__(...)", "as.data.frame", "Extracts a data frame",
  "Misc", ".__length__()", "length", "Returns a custom length, rather the length of the R6 object",
  "Misc", ".__names()__", "names", "Returns names, rather the names of the R6 object",
  "Misc", ".__str(...)__", "str", "Returns a custom strcuture, rather the structure of the R6 object"
)

tbl <-
  gt::gt(
    R6methods,
    groupname_col = "Group"
  ) %>% 
    gt::cols_width(
      vars(`Dot-Dunder Method`, `S3 Method`) ~ gt::px(225),
      vars(Description) ~ gt::px(300)
    )
gt::gtsave(
  data = tbl,
  filename = "supported_methods.png",
  path = "inst/"
)
```

### Iteration

There is one other special method, `.__getitem__(...)`, which, if defined, will allow you easily turn your `R6` class into an iterator. You must also define the `._length__()` method. You can check if your `R6` instance is iterable by calling `R6methods::is.iterable(myR6class)`.

If your `R6` instance is iterable, you can call `R6methods::iter(myR6instance)` to turn your instance into a `coro` iterator. The returned object is a `generator` (i.e. function factory). Calling this `generator` will produce an iterator that iterates the `length()` of your `R6` instance, producing batches of data according to the `.__getitem__()` method.

```{r iterator}
myClass <- R6::R6Class(
  classname = "myClass",
  public = list(
    data = head(mtcars),
    .__length__ = function() {
      nrow(self$data)
    },
    .__getitem__ = function(...) {
      self$data[..1, , drop = FALSE]
    }
  )
)

x <- myClass$new() # Create an instance
gen <- iter(x) # Create the generator
ii <- gen() # Create the iterator

# Collect a single batch
coro::collect(ii, 1)

# Collect two batches
coro::collect(ii, 2)

# Collect remaining batches
coro::collect(ii)
coro::collect(ii) # no batches left
coro::is_exhausted(ii()) # iterator is exhausted

# Create new iterator
ii2 <- gen()

# Loop over batches
coro::loop(for (i in ii2) {
  print(i)
})
```

## Code of Conduct

Please note that the R6methods project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
